---
title: "Post analysis of SSR simulation results"
output: html_document
---


## Introduction

I've created 10,000 simulations of populations undergoing different levels of
sexual reproduction and analyzed the index of association along with genotypic
and allelic diversity metrics, including an analysis of the contribution of each
locus to the genotypic diversity. This will serve as a report for the 
simulations.

## Setup

```{r, required_packages, message = FALSE}
library('zksimanalysis') # Main package for analysis
library('magrittr')      # For the %T>%
library('viridis')       # Excellent color scaling
library('dplyr')         # Magic data manipulation
library('tidyr')         # more magic
library('purrr')         # magic dealing with lists
library('ggplot2')       # magic plotting action!
```

This here will load in all of the results of the analyses

```{r, load_data, cache = TRUE, results = "hide"}
# Here we have to filter out the stored genetic data
datfiles  <- dir("../simulation_results/rda_files/", full.names = TRUE) %>% 
  grep("feather.rda$", ., value = TRUE)
divfiles  <- dir("../simulation_results/diversity_rda_files/", full.names = TRUE)
locfiles  <- dir("../simulation_results/locus_rda_files/", full.names = TRUE)
contfiles <- dir("../simulation_results/contrib_rda_files/", full.names = TRUE)

for (i in datfiles){
  cat(i, "\n")
  load(i)
}

for (i in divfiles){
  cat(i, "\n")
  load(i)
}

for (i in locfiles){
  cat(i, "\n")
  load(i)
}

for (i in contfiles){
  cat(i, "\n")
  load(i)
}
```

```{r, data_tidy, cache = TRUE}
ex_run  <- "twenty_loci([0-9]+?)/"
ex_seed <- "seed_([0-9]+?)_"
ex_sex  <- "sex_([0-9.]+?)_"
ex_gen  <- "gen_([0-9]+?)_"
ex_rep  <- "rep_([0-9]+?).pop"
ex_samp <- "_sam_([0-9]+?)$"

datnames <- grep("^X.+?twenty.+?feather$", ls(), value = TRUE)
divnames <- grep("^X.+?twenty.+?feather.divtable.rda$", ls(), value = TRUE)
locnames <- grep("^X.+?twenty.+?feather.locustable.rda$", ls(), value = TRUE)
contnames <- grep("^X.+?twenty.+?feather.contrib.rda$", ls(), value = TRUE)

datalist <- datnames %>%
  lapply(get) %>%
  setNames(datnames) %>%
  bind_rows(.id = "source") %>%
  select(Ia:pop, source) %>%
  extract(pop, c("run", "seed", "sexrate", "gen", "rep", "sample"),
          paste0(ex_run, ex_seed, ex_sex, ex_gen, ex_rep, ex_samp),
          remove = FALSE) %>%
  mutate(sample = factor(sample, unique(sample)))

divlist <- divnames %>%
  lapply(get) %>%
  setNames(divnames) %>%
  bind_rows()

loclist <- locnames %>%
  lapply(get) %>%
  setNames(locnames) %>%
  bind_rows()

contlist <- contnames %>%
  lapply(get) %>%
  setNames(contnames) %>%
  bind_rows()




CF <- function(NMLG, sample){
  sample <- as.integer(as.character(sample))
  cf <- 1 - (NMLG/sample)
  cf <- (sample/(sample - 1))*cf
  return(cf)
}


datalist <- full_join(datalist, divlist, by = "pop") %>%
  full_join(loclist, by = "pop") %>%
  full_join(contlist, by = "pop") %>%
  mutate(CF = CF(NMLG, sample))

vals <- datalist %>%
  mutate(mean.rd = vapply(samples.rd, mean, numeric(1), na.rm = TRUE)) %>%
  mutate(sd.rd   = vapply(samples.rd, sd, numeric(1), na.rm = TRUE)) %>%
  mutate(var.rd  = vapply(samples.rd, var, numeric(1), na.rm = TRUE)) %>%
  mutate(min.rd  = vapply(samples.rd, min, numeric(1), na.rm = TRUE)) %>%
  mutate(max.rd  = vapply(samples.rd, max, numeric(1), na.rm = TRUE)) %>%
  # mutate(mean.rdcc = vapply(samples.rdcc, mean, numeric(1), na.rm = TRUE)) %>%
  # mutate(sd.rdcc   = vapply(samples.rdcc, sd, numeric(1), na.rm = TRUE)) %>%
  # mutate(var.rdcc  = vapply(samples.rdcc, var, numeric(1), na.rm = TRUE)) %>%
  # mutate(min.rdcc  = vapply(samples.rdcc, min, numeric(1), na.rm = TRUE)) %>%
  # mutate(max.rdcc  = vapply(samples.rdcc, max, numeric(1), na.rm = TRUE)) %>%
  select(-starts_with("samples"))
```


```{r, rd_sexrate}
ggplot(vals, aes(x = sexrate, y = rbarD, fill = sample)) +
  geom_boxplot()
```


```{r, rd_pval}
ggplot(vals, aes(x = sexrate, y = p.rD, fill = sample)) +
  geom_boxplot() +
  scale_y_log10(breaks = c(0.01, 0.1, 1),
                minor_breaks = c((1:9)*0.001, (2:9)*0.01, (2:9)*0.1)) +
  theme_bw()
```

```{r, rd_sexrate_pval}
ggplot(vals, aes(x = sexrate, y = rbarD, color = log(p.rD))) +
  geom_jitter(alpha = 0.25, height = 0) +
  geom_boxplot(color = "black", notch = TRUE, width = 0.5, alpha = 0.25) +
  scale_color_viridis(option = "viridis", breaks = log(c(0.005, 0.01, 0.025, 0.05, 0.1)), labels = c(0.005, 0.01, 0.025, 0.05, 0.1)) +
  facet_wrap(~sample, nrow = 1)
```


```{r, clone-correction}
ggplot(vals, aes(x = rbarD, y = rbarDcc, color = log(p.rD))) +
  geom_point(alpha = 0.25) +
  scale_color_viridis(option = "viridis", breaks = log(c(0.005, 0.01, 0.025, 0.05, 0.1)), labels = c(0.005, 0.01, 0.025, 0.05, 0.1)) +
  geom_abline(slope = 1) +
  facet_grid(sample~sexrate) +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  xlab(expression(bar(r)[d])) +
  ylab(expression(paste(bar(r)[d], " (clone-corrected)")))
```


```{r, rd-cc-sexrate}
ggplot(vals, aes(x = sexrate, y = rbarD - rbarDcc, color = log(p.rD))) +
  geom_point(alpha = 0.25, position = "jitter") +
  geom_boxplot(color = "black", notch = TRUE, width = 0.5, alpha = 0.25) +
  scale_color_viridis(option = "viridis", breaks = log(c(0.005, 0.01, 0.025, 0.05, 0.1)), labels = c(0.005, 0.01, 0.025, 0.05, 0.1)) +
  facet_wrap(~sample, nrow = 1)
```




```{r, Hexp_rbarD}
ggplot(vals, aes(x = Hexp, y = rbarD, color = log(p.rD))) +
  geom_point(alpha = 0.25) +
  scale_color_viridis()
```

```{r, Hexp_rbarD_facet}
ggplot(vals, aes(x = Hexp, y = rbarD, color = log(p.rD))) +
  geom_point(alpha = 0.25) +
  scale_color_viridis() +
  facet_grid(sexrate~sample)
```

```{r, gE5_lE5}
ggplot(vals, aes(x = Evenness, y = E.5.x, color = rbarD)) +
  geom_point(alpha = 0.25) +
  scale_color_viridis() +
  facet_grid(sexrate~sample)
```

```{r}
ggplot(filter(vals, p.rD > 0.95), aes(x = Evenness, y = rbarD, color = E.5.x)) +
  geom_point(alpha = 0.25) +
  scale_color_viridis(option = "magma", direction = -1) +
  facet_grid(sexrate~sample) +
  theme_dark()
```

```{r, grabbag}
ggplot(vals, aes(x = sexrate, y = E.5.x, fill = sample)) +
  geom_boxplot()
ggplot(vals, aes(x = sexrate, y = H.x, fill = sample)) +
  geom_boxplot()
ggplot(vals, aes(x = sexrate, y = B, fill = sample)) +
  geom_boxplot()
```


```{r, CF_B}
ggplot(vals, aes(x = CF, y = B, color = sample)) +
  geom_point() +
  facet_wrap(~sample)
```

