---
title: "ROC Calculations"
output: html_notebook
---

```{r, required_packages}
library('zksimanalysis')
```

This here will load in all of the results of the analyses. These were previously
processed in the file `ssr_data_cleaning.Rmd`. The name of the data was called
"vals".

```{r, load_data, cache = TRUE, results = "hide"}
system.time({
res        <- load(here::here("data", "processed_results.rda"))
vals_by_cc <- get(res) %>% 
  gather(clone_correction, pval, p.rD, p.rDcc) %>%
  mutate(clone_correction = ifelse(clone_correction == "p.rD", "wd", "cc"))
genomic_res <- load(here::here("data", "genomic_data.rda"))
genomic_dat <- get(genomic_res[genomic_res == "original_redux"])
rm(list = c(res, genomic_res))
})
vals_by_cc
genomic_dat
```

## Data Analysis

### Overall ROC Calculation

```{r, generate_sexrate}
sex <- unique(vals_by_cc$sexrate)
sex <- sex[!sex %in% "1.0000"]
alpha <- seq(0, 1, by = 0.01)
```

The main ROC calculation is handled by the function `roc()`. This function
calculates ROC and the variance for a single level of $\alpha$ over a given
grouping. It assumes the column for the factor dividing the true positive and
false positive is called "sexrate". This will calculate ROC and the variance
for each level of the grouping (which must be at least sexrate). 

Since we want create a curve, here we are mapping the `roc()` function over the 
different levels of $\alpha$. We do this for both the regular p-value and clone 
corrected p-value. We then do this for each rate of sexual reproduction 
independently. To do this, we have to use nested `map()` functions. The first 
`map_df()` will map the second map function over the values of non-random
mating, `map2_df()`, which will map the `roc()` function over both the incoming
value of non-random mating (coming in as `.x`) and all values of $\alpha$
(represented as `.y`).

We do this for both full and clone-corrected data.

```{r, roc_overall, cache = TRUE}
Sys.time()
total_roc <- map_df(alpha, roc, 
                    vals_by_cc %>% mutate(sexrate = ifelse(sexrate == "1.0000", "R", "NR")),
                    compare = c("NR", "R"), stat = "pval", count.na = TRUE,
                    group = c("mutation_rate", "sexrate", "sample", "clone_correction"))
total_roc

Sys.time()
roc_overall <- sex %>% 
  map_df(.f = map2_df, alpha, 
         ~roc(.y, vals_by_cc, compare = c(.x, "1.0000"), count.na = TRUE, 
              group = c("mutation_rate", "sexrate", "sample", "clone_correction"), 
              stat = "pval"
             )
         ) 
roc_overall

Sys.time()
```


```{r, roc_by_seed, cache = TRUE}
seedgroup <- c("mutation_rate", "sexrate", "run", "seed", "sample", "clone_correction")
Sys.time()
roc_by_seed <- sex %>% 
  map_df(.f = map2_df, alpha, 
         ~roc(.y, vals_by_cc, compare = c(.x, "1.0000"), count.na = TRUE, 
              group = seedgroup, stat = "pval"
             )
         ) 
roc_by_seed
Sys.time()
```

### ROC considering Allelic evenness

It has been observed that the values for $\bar{r}_d$ can be non-significant and
even sharply negative for purely clonal populations. We also noticed that the
allelic evenness for these populations is quite high, which is a sign of fixed
heterozygosity in several loci. We will use this to "flip" a decision in the
ROC analysis to see if this affects clonal inference. To do this, we will add a
condidtional statement to the ROC analysis and ask if p <= alpha OR Ea > 0.85.
This way, we will be able to correct the p-values that are non-siginficant.

```{r roc_overall_ea, cache = TRUE}
vals_by_cc_ea <- vals_by_cc %>% mutate(ea = Evennesscc > 0.85)
Sys.time()
total_roc_ea <- map_df(alpha, roc, 
                    vals_by_cc_ea %>% mutate(sexrate = ifelse(sexrate == "1.0000", "R", "NR")) , 
                    flip = "ea",
                    compare = c("NR", "R"), stat = "pval", count.na = TRUE,
                    group = c("mutation_rate", "sexrate", "sample", "clone_correction"))
total_roc_ea


Sys.time()
roc_overall_ea <- sex %>% 
  map_df(.f = map2_df, alpha, 
         ~roc(.y, vals_by_cc_ea, flip = "ea", compare = c(.x, "1.0000"), count.na = TRUE, 
              group = c("mutation_rate", "sexrate", "sample", "clone_correction"), 
              stat = "pval"
             )
         ) 
roc_overall_ea
Sys.time()
```

```{r roc_by_seed_ea, cache = TRUE}
seedgroup <- c("mutation_rate", "sexrate", "run", "seed", "sample", "clone_correction")
Sys.time()
roc_by_seed_ea <- sex %>% 
  map_df(.f = map2_df, alpha, 
         ~roc(.y, vals_by_cc_ea, compare = c(.x, "1.0000"), count.na = TRUE, 
              group = seedgroup, stat = "pval", flip = "ea"
             )
         ) 
roc_by_seed_ea
Sys.time()
```


### AURC

Here we calculate the Area Under the ROC Curve using the function `auc()` from
the package *flux*.

```{r, AURC_calc, cache = TRUE}
# Calculate ROC over all seeds, produce points

AURC_calc <- . %>%
  summarize(AURC = auc(`False Positive`, `True Positive`)) %>%
  ungroup() %>% 
  mutate(sample = factor(paste("n =", sample), paste("n =", c(10, 25, 50, 100))))

Sys.time()
AURC_overall <- roc_overall %>% 
  group_by(sexrate, sample, mutation_rate, clone_correction) %>%
  AURC_calc

Sys.time()
# Calculate ROC for each seed separately, produce distributions
AURC_by_seed <- roc_by_seed %>% 
  group_by(run, seed, sexrate, sample, mutation_rate, clone_correction) %>%
  AURC_calc

Sys.time()
AURC_overall_ea <- roc_overall_ea %>% 
  group_by(sexrate, sample, mutation_rate, clone_correction) %>%
  AURC_calc

Sys.time()
# Calculate ROC for each seed separately, produce distributions
AURC_by_seed_ea <- roc_by_seed_ea %>% 
  group_by(run, seed, sexrate, sample, mutation_rate, clone_correction) %>%
  AURC_calc

Sys.time()
```

## Genomic Data

Analyzing the genomic ddata is a bit simpler because we don't have to deal with
specifics such as mutation rate or clone correction (yet). I'm following the
same procedures as above.

```{r genomic_roc, cache = TRUE, dependson = "load_data"}
Sys.time()
g_total_roc <- map_df(alpha, roc, 
                      genomic_dat %>% mutate(sexrate = ifelse(sexrate == "1.0000", "R", "NR")),
                      compare = c("NR", "R"), stat = "p.rD", count.na = TRUE,
                      group = c("sexrate", "sample"))
g_total_roc

Sys.time()
g_roc_overall <- sex %>% 
  map_df(.f = map2_df, alpha, 
         ~roc(.y, genomic_dat, compare = c(.x, "1.0000"), count.na = TRUE, 
              group = c("sexrate", "sample"), 
              stat = "p.rD"
             )
         ) 
g_roc_overall
Sys.time()
g_roc_by_seed <- sex %>% 
  map_df(.f = map2_df, alpha, 
         ~roc(.y, genomic_dat, compare = c(.x, "1.0000"), count.na = TRUE, 
              group = c("sexrate", "sample", "run", "seed"), 
              stat = "p.rD"
             )
         ) 
g_roc_overall
Sys.time()

```

```{r genomic_AURC, cache = TRUE, dependson = "load_data"}
Sys.time()
g_AURC_overall <- g_roc_overall %>% 
  group_by(sexrate, sample) %>%
  AURC_calc
g_AURC_overall
Sys.time()
g_AURC_by_seed <- g_roc_by_seed %>% 
  group_by(sexrate, sample, run, seed) %>%
  AURC_calc
Sys.time()
g_AURC_overall
g_AURC_by_seed
```

## Saving Data

```{r save_ROC_data}
save(total_roc, roc_overall, roc_by_seed, AURC_overall, AURC_by_seed,
     total_roc_ea, roc_overall_ea, roc_by_seed_ea, AURC_overall_ea, AURC_by_seed_ea,
     g_total_roc, g_roc_overall, g_roc_by_seed, g_AURC_overall, g_AURC_by_seed,
     file = here::here("data/ROC_data.rda"))
```



## Session Information

```{r session_info}
options(width = 100)
devtools::session_info()
```
